Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-11-06 paper/article.tex
Paper progress

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. This design aims to solve two
fundamental challenges in using neural stimulation to improve external task performance.
First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
the need for a stimulation agent to account for-far distant effects of the stimulation it
applies. Second, with neurologically complex tasks, such as grasping, the mapping between
neural activity as-measured and the external task cannot be determined \textit{a priori}.

Our co-processor attempts to solve this with a pair of artificial neural networks:
\begin{itemize}
	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
	      between the stimulation, neural dynamics, and external task. It's purpose is for training the second network.
	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
	      data from external sensors, to stimulation parameters.
\end{itemize}

We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
improve task performance. By continually training both, they adapt to the brain as it changes,
effectively causing brain-stimulator co-adaptation. The EN is a tool for training the CPN,
giving us a way to back-propagate task error to the CPN. It outputs task-relevant
metrics - a prediction of muscle velocities in our case - given measurements of neural
activity, and the stimulation parameters output from the CPN. If the EN is trained in a
particular way, and to a sufficient level of precision, it can be used to overcome the
challenge of not knowing a mapping between stimulation and the task. When training the
CPN, we in-effect treat the EN's output as the true task performance, or a related
metric, and then backpropagate the loss defined in terms of that metric. See Figs X, Y.
We will illustrate the details of the training algorithm below.

@@

There are some key differences between that approach and the EN architecture we
present here. Our EN's purpose is simply to train the CPN, so we pick an
architecture we find to be best suited to that job. First, our EN architecture
is not linear, through the use of a $tanh$ activation function. That allows us
to escape some of the limits in representational power of linear models.
Second, LSTM cells are designed to better capture long-running network
dynamics, and in our case we found that they were crucial for that reason.
Compare that to the simpler cells commonly used in vanilla RNNs, which
yield functionally the same approach as the linear state space model, less
the use of a nonlinear activation function.

// TODO? Try having no nonlinearity?

\subsection{Simulation Overview}
* Proof-of-concept by way of simulation.
** Simulation allows for rapid, cheap exploration of learning algorithms.
** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
** Cite Michaels, and those upstream from him (Susillo?) re: RNNs. Networks have internal dynamics that result from both inputs and internal connectivity. Perturbations at t=0 affect outputs at t=N.
** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
  simulation to develop and evaluate a neural controller. Use of lesion model.

>>>>

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. This design aims to solve two
fundamental challenges in using neural stimulation to improve external task performance.
First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
the need for a stimulation agent to account for far-distant effects of the stimulation it
applies. Second, with neurologically complex tasks, such as grasping, the mapping between
neural activity as-measured and the external task cannot be determined \textit{a priori}.
As a result, the co-processor must somehow learn what stimulation is appropriate for
aiding the user in the external task they are attempting to perform.

Our co-processor attempts to solve these with a pair of artificial neural networks:
\begin{itemize}
	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
	      between the stimulation, neural dynamics, and external task. It's purpose is for training the second network.
	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
	      data from external sensors, to stimulation parameters.
\end{itemize}

We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
improve task performance. By continually training both, they adapt to the brain as it changes,
effectively allowing for brain-stimulator co-adaptation. The EN is a tool for training the CPN,
giving us a way to back-propagate task error to the CPN. It outputs task-relevant
metrics - a prediction of muscle velocities in our case - given measurements of neural
activity, and the stimulation parameters output from the CPN. If the EN is trained in a
particular way, and to a sufficient level of precision, it can be used as a function
approximator relating stimulation and a task, and do so in a way that allows us to train
the CPN with it. When training the CPN, we in-effect treat the EN's output as the true
task performance, or a related metric, and then backpropagate the loss defined in terms
of that metric in order to train the CPN. See Fig. \ref{fig:weill}. We will illustrate the details
of the training algorithm below.

@@

There are some key differences between that approach and the EN architecture we
present here. Our EN's purpose is simply to train the CPN, so we pick an
architecture we find to be best suited to that job. First, our EN architecture
is not linear, through the use of a $tanh$ activation function. That allows us
to escape some of the limits in representational power of linear models, though
we didn't deeply test a strictly linear approach.
Second, LSTM cells are designed to better capture long-running network
dynamics, and in our case we found that they were crucial for that reason.
Compare that to the simpler cells commonly used in vanilla RNNs, which
yield functionally the same approach as the linear state space model, less
the use of a nonlinear activation function.

\subsection{Simulation Overview}
We demonstrate our approach here using a simulated grasping circuit. A detailed
simulation such as this allows us to explore some of the critical details in
how to train a co-processor of this design. By first doing such exploration in
simulation, we are able to rapidly and cheaply iterate on our design, prior to
any \textit{in vivo} experiments. In order for the simulation to be
admissible, we need to ensure it has some properties that allow it to strongly
indicate if our design is improving in a direction that will later allow for
real deployments. Otherwise, our design may be adapting to the pecularities of
the simulation, witheout becoming more useful.

** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
** Cite Michaels, and those upstream from him (Susillo?) re: RNNs. Networks have internal dynamics that result from both inputs and internal connectivity. Perturbations at t=0 affect outputs at t=N.
** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
  simulation to develop and evaluate a neural controller. Use of lesion model.
